Assessing primitives performance\\ on multi-stage execution \todo{modify maybe}

### 2018-06-28 paper.tex
comply wih template workshop. Will have to modify reference as well

%%%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[10pt, sigplan]{acmart}
%%\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[sigplan,10pt]{acmart}\settopmatter{}

%% Conference information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
%\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
%\acmYear{2017}
%\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
%\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style

%\bibliographystyle{ACM-Reference-Format}

%% Citation style
%\citestyle{acmauthoryear}  %% For author/year citations
\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format must update the
%% '\documentclass' and topmatter commands above; see
%% 'acmart-pacmpl-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%% Title information
\title[Assessing primitives performance on multi-stage execution]{Assessing primitives performance\\ on multi-stage execution \todo{modify maybe}}

%% Author with single affiliation.
\author{Sophie Kaleba}
                                        %% can be repeated if necessary
\affiliation{
  %\position{Position1}
  \department{RMoD}              %% \department is recommended
  \institution{Inria}            %% \institution is required
 % \streetaddress{Street1 Address1}
  \city{Lille}
  %\state{France}
  %\postcode{Post-Code1}
  \country{France}                    %% \country is recommended
}
\email{sophie.kaleba@etudiant.univ-lille1.fr}          %% \email is recommended

%% Author with two affiliations and emails.
\author{Cl\'ement B\'era}
\affiliation{
  % \position{}
	\department{Software Languages Lab}              %% \department is recommended
	\institution{Vrije Universiteit Brussel}            %% \institution is required
	\city{Brussel}
  % \state{}
  % \postcode{}
	\country{Belgium}                    %% \country is recommended
}
\email{clement.bera@vub.ac.be}          %% \email is recommended

%% Author with two affiliations and emails.
\author{St\'ephane Ducasse}
\affiliation{
 % \position{Position2b}
  \department{RMoD}             %% \department is recommended
  \institution{Inria}           %% \institution is required
  %\streetaddress{Street3b Address2b}
  \city{Lille}
  %\state{State2b}
  %\postcode{Post-Code2b}
  \country{France}                   %% \country is recommended
}
\email{stephane.ducasse@inria.fr}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}

@@

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10011007.10011006.10011008</concept_id>
%<concept_desc>Software and its engineering~General programming languages</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%<concept>
%<concept_id>10003456.10003457.10003521.10003525</concept_id>
%<concept_desc>Social and professional topics~History of programming languages</concept_desc>
%<concept_significance>300</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Software and its engineering~General programming languages}
%\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{Just-in-Time compiler, Primitive, Virtual machine, Managed runtime}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

@@

%% Acknowledgments
%%\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
%  This material is based upon work supported by the
%  \grantsponsor{GS100000001}{National Science
%    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
%  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
%  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
%  conclusions or recommendations expressed in this material are those
%  of the author and do not necessarily reflect the views of the
%  National Science Foundation.
%\end{acks}

%% Bibliography
\bibliographystyle{alpha}
\bibliography{sista,rmod,others}

%\vspace{10cm} %Adapt to make vfill work

%% Appendix
%\null \vfill\null %switch column
%
%\appendix
%\section{Smart Syntax code}
%\label{SmartSyntaxAppendix}
%
%Source code of the smart syntax method, which is used to generate the C code for the plugin but also as Smalltalk fall-back code.
%
%\begin{code}
%compare: string1 with: string2 collated: order
%	"Return 1, 2 or 3, if string1 is <, =, or > string2, with the collating order of characters given by the order array."
%
%	| len1 len2 c1 c2 |
%	<primitive: 'primitiveCompareString' module: 'MiscPrimitivePlugin'>
%	<var: #string1 declareC: 'unsigned char *string1'>
%	<var: #string2 declareC: 'unsigned char *string2'>
%	<var: #order declareC: 'unsigned char *order'>
%
%	len1 := string1 size.
%	len2 := string2 size.
%	1 to: (len1 min: len2) do:
%		[:i |
%		c1 := order at: (string1 basicAt: i) + 1.
%		c2 := order at: (string2 basicAt: i) + 1.
%		c1 = c2 ifFalse: 
%			[c1 < c2 ifTrue: [^ 1] ifFalse: [^ 3]]].
%	len1 = len2 ifTrue: [^ 2].
%	len1 < len2 ifTrue: [^ 1] ifFalse: [^ 3].
%\end{code}
%
%\vfill\null %switch column
%\section{Slang Plugin code}
%\label{PluginAppendix}
%
%Source code of the plugin method generating the C code for the Slang plugin code. Most operations needs to go through the interpreter proxy. 
%
%\begin{code}
%primitiveCompareString
%	"ByteString (class) compare: string1 with: string2 collated: order"
%	<export: true>
%	| len1 len2 order string1 string2 |
%
%	<var: 'order' type: #'unsigned char *'>
%	<var: 'string1' type: #'unsigned char *'>
%	<var: 'string2' type: #'unsigned char *'>
%	((interpreterProxy isBytes: (interpreterProxy stackValue: 0))
%	and: [(interpreterProxy isBytes: (interpreterProxy stackValue: 1))
%	and: [interpreterProxy isBytes: (interpreterProxy stackValue: 2)]]) ifFalse:
%		[^interpreterProxy primitiveFailFor: PrimErrBadArgument].
%	string1 := interpreterProxy firstIndexableField: (interpreterProxy stackValue: 2).
%	string2 := interpreterProxy firstIndexableField: (interpreterProxy stackValue: 1).
%	order := interpreterProxy firstIndexableField: (interpreterProxy stackValue: 0).
%	len1 := interpreterProxy sizeOfSTArrayFromCPrimitive: string1.
%	len2 := interpreterProxy sizeOfSTArrayFromCPrimitive: string2.
%	(interpreterProxy sizeOfSTArrayFromCPrimitive: order) < 256 ifTrue:
%		[^interpreterProxy primitiveFailFor: PrimErrBadArgument].
%	0 to: (len1 min: len2) - 1 do: 
%		[ :i | | c1 c2 |
%		c1 := order at: (string1 at: i).
%		c2 := order at: (string2 at: i).
%		c1 = c2 ifFalse:
%			[^interpreterProxy methodReturnInteger: (c1 < c2 ifTrue: [1] ifFalse: [3])]].
%	interpreterProxy methodReturnInteger:
%		(len1 = len2 ifTrue: [2] ifFalse: [len1 < len2 ifTrue: [1] ifFalse: [3]])
%\end{code}
%
%\vfill\null %switch column
%\section{Slang code}
%\label{SlangAppendix}
%
%Source code of the slang method generating the C code for the Slang string comparison primitive. We split it in two methods, and forced inlining at Slang-to-C compilation time to remove the closures while avoiding code duplication.
%
%\begin{code}
%primitiveCompareWith
%	"<string1> primitiveCompareWith: string2 [collated: order] "
%	<export: true>	
%	| string1 string2 order strLength1 strLength2 result |
%	"1 - fetch the parameters from the stack"	
%	argumentCount = 1 ifFalse:
%		[argumentCount ~= 2 ifTrue:
%			[^self primitiveFailFor: PrimErrBadNumArgs].
%			 order := self stackTop.
%			 ((objectMemory isBytes: order)
%			 and: [(objectMemory numBytesOfBytes: order) = 256]) ifFalse:
%				[^self primitiveFailFor: PrimErrBadArgument]].
%	string1 := self stackValue: argumentCount.
%	string2 := self stackValue: argumentCount - 1. 		
%	"2 - check their types - all parameters are ByteObject"
%	((objectMemory isBytes: string1)
%	 and: [objectMemory isBytes: string2]) ifFalse: 
%		[^self primitiveFailFor: PrimErrBadArgument].
%	"3 - compare the strings"	
%	strLength1 := objectMemory numBytesOfBytes: string1.
%	strLength2 := objectMemory numBytesOfBytes: string2.
%	result := order 
%		ifNil: [self rawCompare: string1 length: strLength1 with: string2 length: strLength2 accessBlock:
%				[:str :index | objectMemory fetchByte: index ofObject: str ]]
%		ifNotNil: 
%			[self rawCompare: string1 length: strLength1 with: string2 length: strLength2 accessBlock:
%				[:str :index | objectMemory fetchByte: (objectMemory fetchByte: index ofObject: str) + 1 ofObject: order ]].
%
%	self methodReturnInteger: result
%
%
%rawCompare: string1 length: strLength1 with: string2 length: strLength2 accessBlock: accessBlock
%	| c1 c2 min |
%	<inline: true> "needs to be forced else slang does not inline it by default"
%	min := strLength1 min: strLength2.
%	0 to: min-1 do: 
%		[:i | c1 := accessBlock value: string1 value: i.
%			c2 := accessBlock value: string2 value: i.
%			c1 = c2 ifFalse: [^c1 - c2]].
%	^strLength1 - strLength2
%\end{code}
%
%\vfill\null %switch column
%\section{Cog's RTL code}
%\label{RTLAppendix}
%
%Source code of the method generating the native code of the string comparison primitive on top of Cog's RTL. Method names starting with uppercase (in blue) are directly generating an abstract assembly instruction. Other methods called, starting with gen, generate a sequence of abstract assembly instructions. We duplicated the code checking if each operand is a ByteString and extract its size to generate better native code than doing the two operations separatedly.
%
%\begin{code}
%genPrimitiveStringCompareWith
%	| instr jump jumpAbove jumpIncorrectFormat1 jumpIncorrectFormat2 jumpIncorrectFormat3 jumpIncorrectFormat4 jumpMidFailure jumpSuccess minSizeReg string1CharOrByteSizeReg string2CharOrByteSizeReg string1Reg string2Reg |
%	<var: #jumpIncorrectFormat1 type: #'AbstractInstruction *'>
%	<var: #jumpIncorrectFormat2 type: #'AbstractInstruction *'>
%	<var: #jumpIncorrectFormat3 type: #'AbstractInstruction *'>
%	<var: #jumpIncorrectFormat4 type: #'AbstractInstruction *'>
%	<var: #jumpAbove type: #'AbstractInstruction *'>
%	<var: #jumpSuccess type: #'AbstractInstruction *'>
%	<var: #jump type: #'AbstractInstruction *'>
%	<var: #jumpMidFailure type: #'AbstractInstruction *'>
%	"I redefine those name to ease program comprehension"
%	string1Reg := ReceiverResultReg.
%	string2Reg := Arg0Reg.
%	string1CharOrByteSizeReg := Arg1Reg.
%	string2CharOrByteSizeReg := ClassReg.
%	minSizeReg := SendNumArgsReg.
%	"Load arguments in reg"
%	cogit genLoadArgAtDepth: 0 into: string2Reg.
%	"checks if string1 is a byteobject and get its size in bytes"
%	self genGetFormatOf: string1Reg into: TempReg.
%	cogit CmpCq: objectMemory firstByteFormat R: TempReg.
%	jumpIncorrectFormat1 := cogit JumpLess: 0.
%	cogit CmpCq: objectMemory firstCompiledMethodFormat R: TempReg.
%	jumpIncorrectFormat2 := cogit JumpAboveOrEqual: 0.	
%	self genGetNumSlotsOf: string1Reg into: string1CharOrByteSizeReg.
%	(cogit LogicalShiftLeftCq: objectMemory shiftForWord R: string1CharOrByteSizeReg). 
%	cogit AndCq: objectMemory wordSize - 1 R: TempReg R: TempReg. 
%	cogit SubR: TempReg R: string1CharOrByteSizeReg. 
%	"checks if string2 is a byteobject and get its size in bytes"
%	self genGetFormatOf: string2Reg into: TempReg.
%	cogit CmpCq: objectMemory firstByteFormat R: TempReg.
%	jumpIncorrectFormat3 := cogit JumpLess: 0.
%	cogit CmpCq: objectMemory firstCompiledMethodFormat R: TempReg.
%	jumpIncorrectFormat4 := cogit JumpAboveOrEqual: 0.	
%	self genGetNumSlotsOf: string2Reg into: string2CharOrByteSizeReg.
%	(cogit LogicalShiftLeftCq: objectMemory shiftForWord R: string2CharOrByteSizeReg).
%	cogit AndCq: objectMemory wordSize - 1 R: TempReg R: TempReg.
%	cogit SubR: TempReg R: string2CharOrByteSizeReg.
%	\end{code}
%	\vfill\null %switch column
%	\begin{code}
%	"Type and number of arguments are correct"
%	"Compute the min"	 
%	cogit CmpR: string1CharOrByteSizeReg R: string2CharOrByteSizeReg.
%	jumpAbove := cogit JumpBelow: 0. 
%	cogit MoveR: string1CharOrByteSizeReg R: minSizeReg. 
%	jump := cogit Jump: 0. 
%	jumpAbove jmpTarget: (cogit MoveR: string2CharOrByteSizeReg R: minSizeReg). 
%	jump jmpTarget: (cogit CmpCq: 0 R: minSizeReg). 
%	jumpSuccess := cogit JumpZero: 0. "if one of the string is empty, no need to go through the comparing loop"
%	"Compare the bytes"
%	cogit MoveCq: objectMemory baseHeaderSize  R: TempReg.
%	cogit AddCq: objectMemory baseHeaderSize R: minSizeReg.
%	instr := cogit MoveXbr: TempReg R: string1Reg R: string1CharOrByteSizeReg.
%	cogit MoveXbr: TempReg R: string2Reg R: string2CharOrByteSizeReg.
%	cogit SubR: string2CharOrByteSizeReg R: string1CharOrByteSizeReg. 
%	jumpMidFailure := cogit JumpNonZero: 0. "the 2 compared characters are different, exit the loop"
%	cogit AddCq: 1 R: TempReg.
%	cogit CmpR: TempReg R: minSizeReg. 
%	cogit JumpNonZero: instr.
%	"all bytes from 1 to minSize are equal"
%	self genGetNumBytesOf: string1Reg into: string1CharOrByteSizeReg.
%	self genGetNumBytesOf: string2Reg into: string2CharOrByteSizeReg.
%	jumpSuccess jmpTarget: (cogit SubR: string2CharOrByteSizeReg R: string1CharOrByteSizeReg).
%	jumpMidFailure  jmpTarget: (cogit MoveR: string1CharOrByteSizeReg R: ReceiverResultReg).	
%	self genConvertIntegerToSmallIntegerInReg: ReceiverResultReg.
%	cogit genPrimReturn.
%	jumpIncorrectFormat4 
%		jmpTarget: (jumpIncorrectFormat3 
%			jmpTarget: (jumpIncorrectFormat2 
%				jmpTarget: (jumpIncorrectFormat1 jmpTarget: cogit Label))).
%	^ CompletePrimitive}
%\end{code}

>>>>

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint       Remove this option only once the paper is in final form.
%  9pt           Set paper in  9-point type (instead of default 10-point)
% 11pt           Set paper in 11-point type (instead of default 10-point).
% numbers        Produce numeric citations with natbib (instead of default author/year).
% authorversion  Prepare an author version, with appropriate copyright-space text.

\usepackage{amsmath}

@@

\newcommand{\cL}{{\cal L}}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF'yy}{Month d--d, 20yy, City, ST, Country}
\copyrightyear{20yy}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}\reprintprice{\$15.00}
\copyrightdoi{nnnnnnn.nnnnnnn}

% For compatibility with auto-generated ACM eRights management
% instructions, the following alternate commands are also supported.
%\CopyrightYear{2016}
%\conferenceinfo{CONF'yy,}{Month d--d, 20yy, City, ST, Country}
%\isbn{978-1-nnnn-nnnn-n/yy/mm}\acmPrice{\$15.00}
%\doi{http://dx.doi.org/10.1145/nnnnnnn.nnnnnnn}

% Uncomment the publication rights used.
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}  % default
%\setcopyright{rightsretained}

%\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{Assessing primitives performance on multi-stage execution}   % 'preprint' option specified.

\title{Assessing primitives performance\\ on multi-stage execution \todo{modify maybe}}
%\subtitle{Subtitle Text, if any\titlenote{with optional subtitle note}}

\authorinfo{Sophie Kaleba}
           {RMoD - Inria Lille}
           %{Lille, France}
           {sophie.kaleba@etudiant.univ-lille1.fr}

\authorinfo{Cl\'ement B\'era}
           {Software Languages Lab - Vrije Universiteit Brussel}
           {clement.bera@vub.ac.be}

\authorinfo{{St\'ephane Ducasse}}
           {RMoD - Inria Lille}
           {stepahne.ducasse@inria.fr}

\maketitle

@@

% 2012 ACM Computing Classification System (CSS) concepts
% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10011007.10011006.10011008</concept_id>
%<concept_desc>Software and its engineering~General programming languages</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%<concept>
%<concept_id>10003752.10010124.10010138.10010143</concept_id>
%<concept_desc>Theory of computation~Program analysis</concept_desc>
%<concept_significance>300</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}
%
%\ccsdesc[500]{Software and its engineering~General programming languages}
%\ccsdesc[300]{Theory of computation~Program analysis}
% end generated code

% Legacy 1998 ACM Computing Classification System categories are also
% supported, but not recommended.
%\category{CR-number}{subcategory}{third-level}[fourth-level]
%\category{D.3.0}{Programming Languages}{General}
%\category{F.3.2}{Logics and Meanings of Programs}{Semantics of Programming Languages}[Program analysis]

\keywords
Just-in-Time compiler, Primitive, Virtual machine, Managed runtime

@@

%\appendix
%\section{Appendix Title}
%
%This is the text of the appendix, if you need one.

\acks
\todo{add acknowledgments}
Acknowledgments, if needed.

% The 'abbrvnat' bibliography style is recommended.
\bibliographystyle{alpha}
\bibliography{sista,rmod,others}
%\bibliographystyle{abbrvnat}
%
%% The bibliography should be embedded for final submission.
%
%\begin{thebibliography}{}
%\softraggedright
%
%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%P. Q. Smith, and X. Y. Jones. ...reference text...
%
%\end{thebibliography}
